---
title: "comparison.trip.catchability"
output: html_document
date: "2024-03-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, wdnr.fmdb, wdnr.gis, lubridate, here, lme4, glmmTMB, DHARMa, emmeans, effects)

```

Now comparing statistical modeling outcomes for trip-level angling cpue vs. population density (from PEs) or relative abundance (electrofishing cpe)

```{r}
# walleye data

cpue.pe.wye<-read_csv(here::here("data","indiv.level.cpue.pe.walleye.csv"))%>%
  mutate(ln.pop.density=log(pop.density))%>%
  ungroup()%>%
  mutate(sample.date=ymd(sample.date),
         month=month(sample.date))%>%
  filter(!is.na(pop.density)&!is.na(cpue))%>%
  filter(cpue<20)

ef.cpe.wye<-read_csv(here::here("data","dnr.ef.cpe.walleye.lmb.csv"))%>%
filter(species=="walleye" & gear=="boom_shocker")%>%
  mutate(km.shocked=total_effort*1.61,
         cpe_km=total_catch/km.shocked)%>%
  filter(km.shocked>1)%>%
  group_by(wbic, year)%>%
  summarize(mean_cpe_km=mean(cpe_km),
            sd_cpe_km=sd(cpe_km),
            n_ef_surveys=n())%>%
  ungroup()

# join by year and wbic to walleye cpue

cpue.ef.cpe.wye<-cpue.pe.wye%>%
  dplyr::select(county:shore.length)%>%
  left_join(ef.cpe.wye, by=c("wbic","year"))%>%
  filter(!is.na(mean_cpe_km) & !is.na(cpue))%>%
  mutate(ln.ef.cpe=log(mean_cpe_km))%>%
  filter(cpue<20)%>%
  mutate(month=as.factor(month(sample.date)))
  

# lmb data
cpue.ef.cpe.lmb<-read_csv(here::here("data","indiv.level.cpue.ef.cpe.lmb.csv"))%>%
  dplyr::select(county:shore.length)%>%
  filter(!is.na(mean_cpe_km) & !is.na(cpue))%>%
  mutate(ln.ef.cpe=log(mean_cpe_km))%>%
  filter(cpue<20)%>%
  mutate(month=as.factor(month(sample.date)))
```
Starting with walleye pe data: cpue.pe.wye

```{r}
ggplot(cpue.pe.wye)+
  geom_point(aes(x=pop.density, y=cpue))+
  theme_bw()


ggplot(cpue.pe.wye)+
  geom_point(aes(x=ln.pop.density, y=log(cpue+0.01)))+
  theme_bw()
```


```{r}
mod.null.wye.pe<-glmmTMB(cpue~.pop.density + month, 
                  data=cpue.pe.wye, family=glmmTMB::lognormal(link="log"),
               ziformula=~.pop.density + month)

mod.wbic.wye.pe<-glmmTMB(cpue~.pop.density + month+(1|wbic) , 
                  data=cpue.pe.wye, family=glmmTMB::lognormal(link="log"),
               ziformula=~.pop.density + month+(1|wbic))

mod.wbic.year.wye.pe<-glmmTMB(cpue~.pop.density + month+(1|wbic)+(1|year) , 
                  data=cpue.pe.wye, family=glmmTMB::lognormal(link="log"),
               ziformula=~.pop.density + month+(1|wbic)+(1|year))

mod.wbic.day.wye.pe<-glmmTMB(cpue~.pop.density + month+(1|wbic)+(1|sample.date) , 
                  data=cpue.pe.wye, family=glmmTMB::lognormal(link="log"),
               ziformula=~.pop.density +
                 month+(1|wbic)+(1|sample.date))

mod.wbic.year.day.wye.pe<-glmmTMB(cpue~.pop.density + month+(1|wbic)+(1|year)+(1|sample.date) , 
                  data=cpue.pe.wye, family=glmmTMB::lognormal(link="log"),
               ziformula=~.pop.density +
                 month+(1|wbic)+(1|year)+(1|sample.date))

AIC(mod.null.wye.pe, mod.wbic.wye.pe, mod.wbic.year.wye.pe, mod.wbic.day.wye.pe, mod.wbic.year.day.wye.pe)

sim.resid<-simulateResiduals(mod.wbic.year.day.wye.pe)
plot(sim.resid)

mod.wye.pe<-mod.wbic.year.day.wye.pe
```
```{r}
summary(mod.wye.pe)
performance::r2(mod.wye.pe)
```
woof, incredibly low r2

```{r}
cpue.pe.wye$mod.pred<-predict(mod.wye.pe, type="response")

ggplot(cpue.pe.wye)+
  geom_point(aes(x=pop.density, y=cpue))+
  geom_point(aes(x=pop.density, y=mod.pred), color="red")+
  theme_bw()
```
now walleye ef cpe

```{r}
ggplot(cpue.ef.cpe.wye)+
  geom_point(aes(x=mean_cpe_km, y=cpue))+
  theme_bw()


ggplot(cpue.ef.cpe.wye)+
  geom_point(aes(x=ln.ef.cpe, y=log(cpue+0.01)))+
  theme_bw()
```


```{r}
mod.null.wye.ef.cpe<-glmmTMB(cpue~.ef.cpe + month, 
                  data=cpue.ef.cpe.wye, family=glmmTMB::lognormal(link="log"),
               ziformula=~.ef.cpe + month)

mod.wbic.wye.ef.cpe<-glmmTMB(cpue~.ef.cpe + month+(1|wbic) , 
                  data=cpue.ef.cpe.wye, family=glmmTMB::lognormal(link="log"),
               ziformula=~.ef.cpe + month+(1|wbic))

mod.wbic.year.wye.ef.cpe<-glmmTMB(cpue~.ef.cpe + month+(1|wbic)+(1|year) , 
                  data=cpue.ef.cpe.wye, family=glmmTMB::lognormal(link="log"),
               ziformula=~.ef.cpe + month+(1|wbic)+(1|year))

mod.wbic.day.wye.ef.cpe<-glmmTMB(cpue~.ef.cpe + month+(1|wbic)+(1|sample.date) , 
                  data=cpue.ef.cpe.wye, family=glmmTMB::lognormal(link="log"),
               ziformula=~.ef.cpe +
                 month+(1|wbic)+(1|sample.date))

mod.wbic.year.day.wye.ef.cpe<-glmmTMB(cpue~.ef.cpe + month+(1|wbic)+(1|year)+(1|sample.date) , 
                  data=cpue.ef.cpe.wye, family=glmmTMB::lognormal(link="log"),
               ziformula=~.ef.cpe +
                 month+(1|wbic)+(1|year)+(1|sample.date))

AIC(mod.null.wye.ef.cpe, mod.wbic.wye.ef.cpe, mod.wbic.year.wye.ef.cpe, mod.wbic.day.wye.ef.cpe, mod.wbic.year.day.wye.ef.cpe)

sim.resid<-simulateResiduals(mod.wbic.year.day.wye.ef.cpe)
plot(sim.resid)

mod.wye.ef.cpe<-mod.wbic.year.day.wye.ef.cpe
```
```{r}
summary(mod.wye.ef.cpe)
performance::r2(mod.wye.ef.cpe)
```
low r2, but actually higher than the model fit to PEs

```{r}
cpue.ef.cpe.wye$mod.pred<-predict(mod.wye.ef.cpe, type="response")

ggplot(cpue.ef.cpe.wye)+
  geom_point(aes(x=mean_cpe_km, y=cpue))+
  geom_point(aes(x=mean_cpe_km, y=mod.pred), color="red")+
  theme_bw()
```


And LMB ef cpe


```{r}
ggplot(cpue.ef.cpe.lmb)+
  geom_point(aes(x=mean_cpe_km, y=cpue))+
  theme_bw()


ggplot(cpue.ef.cpe.lmb)+
  geom_point(aes(x=ln.ef.cpe, y=log(cpue+0.01)))+
  theme_bw()
```
Visually, there's already less of a distinct pattern compared to walleye

Convergence issues if I fit the model with month as a covariate. Looking at a histogram of months in the data, I may see the problem: 

```{r}
hist(as.numeric(cpue.ef.cpe.lmb$month))
```
I'm going to adjust the month predictor; 5, 6, 7, 8, 9, 10 will remain as dummy variables; 1, 2, 3, 4, 11, and 12 will be baseline

```{r}
cpue.ef.cpe.lmb$month<-ifelse(cpue.ef.cpe.lmb$month%in%c(1,2,3,4,11,12), 0, cpue.ef.cpe.lmb$month)
hist(cpue.ef.cpe.lmb$month)

cpue.ef.cpe.lmb$month<-as.factor(cpue.ef.cpe.lmb$month)
str(cpue.ef.cpe.lmb$month)
```


```{r}
mod.null.lmb.ef.cpe<-glmmTMB(cpue~.ef.cpe + month, 
                  data=cpue.ef.cpe.lmb, family=glmmTMB::lognormal(link="log"),
               ziformula=~.ef.cpe + month)

mod.wbic.lmb.ef.cpe<-glmmTMB(cpue~.ef.cpe + month+(1|wbic) , 
                  data=cpue.ef.cpe.lmb, family=glmmTMB::lognormal(link="log"),
               ziformula=~.ef.cpe + month+(1|wbic))

mod.wbic.year.lmb.ef.cpe<-glmmTMB(cpue~.ef.cpe + month+(1|wbic)+(1|year) , 
                  data=cpue.ef.cpe.lmb, family=glmmTMB::lognormal(link="log"),
               ziformula=~.ef.cpe + month+(1|wbic)+(1|year))

mod.wbic.day.lmb.ef.cpe<-glmmTMB(cpue~.ef.cpe + month+(1|wbic)+(1|sample.date) , 
                  data=cpue.ef.cpe.lmb, family=glmmTMB::lognormal(link="log"),
               ziformula=~.ef.cpe +
                 month+(1|wbic)+(1|sample.date))

mod.wbic.year.day.lmb.ef.cpe<-glmmTMB(cpue~.ef.cpe + month+(1|wbic)+(1|year)+(1|sample.date) , 
                  data=cpue.ef.cpe.lmb, family=glmmTMB::lognormal(link="log"),
               ziformula=~.ef.cpe +
                 month+(1|wbic)+(1|year)+(1|sample.date))

AIC(mod.null.lmb.ef.cpe, mod.wbic.lmb.ef.cpe, mod.wbic.year.lmb.ef.cpe, mod.wbic.day.lmb.ef.cpe, mod.wbic.year.day.lmb.ef.cpe)

sim.resid<-simulateResiduals(mod.wbic.year.day.lmb.ef.cpe)
plot(sim.resid)

mod.lmb.ef.cpe<-mod.wbic.year.day.lmb.ef.cpe
```
```{r}
summary(mod.lmb.ef.cpe)
performance::r2(mod.lmb.ef.cpe)
```
low, but like ef.cpue.wye, not as low as model using PEs

```{r}
cpue.ef.cpe.lmb$mod.pred<-predict(mod.lmb.ef.cpe, type="response")

ggplot(cpue.ef.cpe.lmb)+
  geom_point(aes(x=mean_cpe_km, y=cpue))+
  geom_point(aes(x=mean_cpe_km, y=mod.pred), color="red")+
  theme_bw()
```

For final comparison, fishscapes data, cpue vs PE and cpue vs ef cpe


```{r}
fishscapes.pe<-read_csv(here::here("data","fishscapes.cpue.pe.ef.cpe.csv"))%>%
  filter(!is.na(pop.density))

fishscapes.ef.cpe<-read_csv(here::here("data","fishscapes.cpue.pe.ef.cpe.csv"))%>%
  filter(!is.na(efCPE))

fishscapes.full<-read_csv(here::here("data","fishscapes.cpue.pe.ef.cpe.csv"))

all_equal(fishscapes.pe, fishscapes.ef.cpe)
# good; just need the one


fishscapes<-fishscapes.pe%>%
  mutate(cpue=anCPUE,
         ln.cpue=log(cpue+0.01),
         month=month(date))%>%
  rename("wbic"=WBIC,
         "ef.cpe"=efCPE)%>%
  mutate(wbic=as.factor(wbic),
         year=as.factor(year),
         month=as.factor(month))
  

ggplot(fishscapes)+
  geom_point(aes(x=pop.density, y=cpue))+
  theme_bw()


ggplot(fishscapes)+
  geom_point(aes(x=ln.pop.density, y=ln.cpue))

hist(as.numeric(fishscapes$month))
```

Fewer months than dnr data; could still use month or day of year

instead of date random effect and month as a fixed effect, adding day of year as quadratic fixed effect.
date will still be tested as a random effect in some models

previous convergence problems (and mismatch with previous models) was caused by year being numeric instead of a factor. silly goose.

```{r}

mod.null.lmb.pe<-glmmTMB(cpue~ln.pop.density + year, 
                  data=fishscapes, family=glmmTMB::lognormal(link="log"),
               ziformula=~ln.pop.density+ sc.doy + sc.doy2 +year)


mod.lakeID.lmb.pe<-glmmTMB(cpue~ln.pop.density + year +(1|lakeID) , 
                  data=fishscapes, family=glmmTMB::lognormal(link="log"),
               ziformula=~ln.pop.density+ sc.doy + sc.doy2 +year+(1|lakeID))




mod.caughtBy.lmb.pe<-glmmTMB(cpue~ln.pop.density + year +(1|caughtBy) , 
                  data=fishscapes, family=glmmTMB::lognormal(link="log"),
               ziformula=~ln.pop.density+ sc.doy + sc.doy2 +year+(1|caughtBy))

mod.date.lmb.pe<-glmmTMB(cpue~ln.pop.density +  year +(1|date) , 
                  data=fishscapes, family=glmmTMB::lognormal(link="log"),
               ziformula=~ln.pop.density+ sc.doy + sc.doy2 +year+(1|date))

mod.lakeID.caughtBy.lmb.pe<-glmmTMB(cpue~ln.pop.density + year+(1|lakeID)+(1|caughtBy), 
                  data=fishscapes, family=glmmTMB::lognormal(link="log"),
               ziformula=~ln.pop.density+ sc.doy + sc.doy2 +year+(1|lakeID)+(1|caughtBy) )

mod.lakeID.date.lmb.pe<-glmmTMB(cpue~ln.pop.density + year+(1|lakeID)+(1|date) , 
                  data=fishscapes, family=glmmTMB::lognormal(link="log"),
               ziformula=~ln.pop.density + sc.doy + sc.doy2+year+(1|lakeID)+(1|date))

mod.caughtBy.date.lmb.pe<-glmmTMB(cpue~ln.pop.density + year+(1|caughtBy)+(1|date) , 
                  data=fishscapes, family=glmmTMB::lognormal(link="log"),
               ziformula=~ln.pop.density + sc.doy + sc.doy2+year+(1|caughtBy)+(1|date))
mod.lakeID.caughtBy.date.lmb.pe<-glmmTMB(cpue~ln.pop.density +year+(1|lakeID)+(1|date)+(1|caughtBy) , 
                  data=fishscapes, family=glmmTMB::lognormal(link="log"),
               ziformula=~ln.pop.density+ sc.doy + sc.doy2+year+(1|lakeID)+(1|date)+(1|caughtBy) )

AIC(mod.null.lmb.pe, mod.lakeID.lmb.pe, mod.caughtBy.lmb.pe, mod.date.lmb.pe, mod.lakeID.caughtBy.lmb.pe, mod.lakeID.date.lmb.pe, mod.caughtBy.date.lmb.pe, mod.lakeID.caughtBy.date.lmb.pe )

sim.resid<-simulateResiduals(mod.caughtBy.date.lmb.pe)
plot(sim.resid)

mod.fs.pe<-mod.caughtBy.date.lmb.pe
```

Interesting; caughtBy and date random effects are best fits; does not include wbic!

```{r}
summary(mod.fs.pe)
performance::r2(mod.fs.pe)
```
higher, but not incredible r2. 

```{r}
fishscapes$mod.pred<-predict(mod.fs.pe, type="response")

ggplot(fishscapes)+
  geom_point(aes(x=pop.density, y=cpue))+
  geom_point(aes(x=pop.density, y=mod.pred), color="red")+
  theme_bw()
```
Repeat, but with ef cpe

```{r}
ggplot(fishscapes)+
  geom_point(aes(x=ef.cpe, y=cpue))+
  theme_bw()


ggplot(fishscapes)+
  geom_point(aes(x=ln.ef.cpe, y=ln.cpue))

```

Fewer months than dnr data; could still use month or day of year

instead of date random effect and month as a fixed effect, adding day of year as quadratic fixed effect.
date will still be tested as a random effect in some models

Moving sc.doy out of conditional model, but keeping it in zero inflated model. (problems with convergence, consistently no effect in conditional model)
```{r}

mod.null.lmb.ef.cpe<-glmmTMB(cpue~ln.ef.cpe  + year, 
                  data=fishscapes, family=glmmTMB::lognormal(link="log"),
               ziformula=~ln.ef.cpe+ sc.doy + sc.doy2 +year)

mod.lakeID.lmb.ef.cpe<-glmmTMB(cpue~ln.ef.cpe + year +(1|lakeID) , 
                  data=fishscapes, family=glmmTMB::lognormal(link="log"),
               ziformula=~ln.ef.cpe+ sc.doy + sc.doy2 +year+(1|lakeID))
# convergence problem
mod.caughtBy.lmb.ef.cpe<-glmmTMB(cpue~ln.ef.cpe +  year +(1|caughtBy) , 
                  data=fishscapes, family=glmmTMB::lognormal(link="log"),
               ziformula=~ln.ef.cpe+ sc.doy + sc.doy2 +year+(1|caughtBy))

mod.date.lmb.ef.cpe<-glmmTMB(cpue~ln.ef.cpe +  year +(1|date) , 
                  data=fishscapes, family=glmmTMB::lognormal(link="log"),
               ziformula=~ln.ef.cpe+ sc.doy + sc.doy2 +year+(1|date))

mod.lakeID.caughtBy.lmb.ef.cpe<-glmmTMB(cpue~ln.ef.cpe + year+(1|lakeID)+(1|caughtBy), 
                  data=fishscapes, family=glmmTMB::lognormal(link="log"),
               ziformula=~ln.ef.cpe+ sc.doy + sc.doy2 +year+(1|lakeID)+(1|caughtBy) )

mod.lakeID.date.lmb.ef.cpe<-glmmTMB(cpue~ln.ef.cpe +year+(1|lakeID)+(1|date) , 
                  data=fishscapes, family=glmmTMB::lognormal(link="log"),
               ziformula=~ln.ef.cpe+ sc.doy + sc.doy2 +year+(1|lakeID)+(1|date))

mod.caughtBy.date.lmb.ef.cpe<-glmmTMB(cpue~ln.ef.cpe +year+(1|caughtBy)+(1|date) , 
                  data=fishscapes, family=glmmTMB::lognormal(link="log"),
               ziformula=~ln.ef.cpe + sc.doy + sc.doy2+year+(1|caughtBy)+(1|date))


mod.lakeID.caughtBy.date.lmb.ef.cpe<-glmmTMB(cpue~ln.ef.cpe + year+(1|lakeID)+(1|date)+(1|caughtBy) , 
                  data=fishscapes, family=glmmTMB::lognormal(link="log"),
               ziformula=~ln.ef.cpe+ sc.doy + sc.doy2 +year+(1|lakeID)+(1|date)+(1|caughtBy) )

AIC(mod.null.lmb.ef.cpe, mod.lakeID.lmb.ef.cpe, mod.caughtBy.lmb.ef.cpe, mod.date.lmb.ef.cpe, mod.lakeID.caughtBy.lmb.ef.cpe, mod.lakeID.date.lmb.ef.cpe, mod.caughtBy.date.lmb.ef.cpe, mod.lakeID.caughtBy.date.lmb.ef.cpe )

sim.resid<-simulateResiduals(mod.lakeID.caughtBy.date.lmb.ef.cpe)
plot(sim.resid)

mod.fs.ef.cpe<-mod.lakeID.caughtBy.date.lmb.ef.cpe
```


```{r}
summary(mod.fs.ef.cpe)
performance::r2(mod.fs.ef.cpe)

performance::r2_zeroinflated(mod.fs.ef.cpe)
```
ef.cpe not a significant predictor. (replaced by wbic?)

looking at predictions (deterministic--no error term) for catch rates. have to model separately zero and non-zero
```{r}
fishscapes$mod.pred<-predict(mod.fs.ef.cpe, type="response")

fishscapes$zero.pred<-predict(mod.fs.ef.cpe, type="zprob")
fishscapes$probCatchFish<-1-fishscapes$zero.pred

fishscapes$caughtFishSim<-rep(NA)


for(i in 1:nrow(fishscapes)){
fishscapes$caughtFishSim[i]<-sample(c(0,1), 1, prob=c(fishscapes$zero.pred[i], fishscapes$probCatchFish[i]))
}

fs.zeros<-filter(fishscapes, caughtFishSim==0)
fs.ones<-filter(fishscapes, caughtFishSim==1)

fs.ones$pred.cpue<-predict(mod.fs.ef.cpe, newdata=fs.ones, type="response")
fs.zeros$pred.cpue<-rep(0)

fs.pred<-rbind.data.frame(fs.ones, fs.zeros)


ggplot(fs.pred)+
  geom_point(aes(x=ef.cpe, y=cpue))+
  geom_point(aes(x=ef.cpe, y=pred.cpue), color="red", alpha=0.5)+
  theme_bw()

ggplot(fs.pred)+
  geom_histogram(aes(cpue), color="gray")+
  geom_histogram(aes(pred.cpue), fill="red", alpha=0.5)

ggplot(fs.pred)+
  geom_point(aes(x=cpue, y=pred.cpue))+
  geom_abline(slope=1, intercept=0, linetype="dashed")+
  theme_bw()
```
Looking at random effects: 

```{r}
rand.cpue<-ranef(mod.fs.ef.cpe)
rand.zero<-ranef(mod.fs.ef.cpe)
```




More stuff interrogating model: predictions and r2 without angler id, full fishscapes dataset without population estimates or electrofishign cpe (didn't work well--residuals were crazy. Even if it doesn't have a significant effect, having some metric of fish population is important)
--------------------------------------------------------------------------------
Look at the next best fitting model that does not include angler effects

```{r}
summary(mod.date.lmb.ef.cpe)
performance::r2(mod.date.lmb.ef.cpe)
```
```{r}
fishscapes.ef.cpe$pred.date<-predict(mod.date.lmb.ef.cpe, type="response")
ggplot(fishscapes.ef.cpe)+
  geom_point(aes(x=ln.ef.cpe, y=anCPUE))+
  geom_point(aes(x=ln.ef.cpe, y=pred.date), color="red")
```









What if I fit the full dataset, avoiding population density entirely? 

one lake doesn't have a WBIC, so using lakeID instead

```{r}
fishscapes.full<-fishscapes.full%>%
  mutate(year=as.factor(year),
         cpue=anCPUE,
         wbic=as.factor(WBIC),
         lakeID=as.factor(lakeID),
         date=as.factor(date))
  

mod.null.lmb<-glmmTMB(cpue~ sc.doy + sc.doy2 + year, 
                  data=fishscapes.full, family=glmmTMB::lognormal(link="log"),
               ziformula=~ sc.doy + sc.doy2 +year)

mod.lakeID.lmb<-glmmTMB(cpue~ sc.doy + sc.doy2+ year +(1|lakeID) , 
                  data=fishscapes.full, family=glmmTMB::lognormal(link="log"),
               ziformula=~ sc.doy + sc.doy2 +year+(1|lakeID))

mod.caughtBy.lmb<-glmmTMB(cpue~ sc.doy + sc.doy2+ year +(1|caughtBy) , 
                  data=fishscapes.full, family=glmmTMB::lognormal(link="log"),
               ziformula=~ sc.doy + sc.doy2 +year+(1|caughtBy))

mod.date.lmb<-glmmTMB(cpue~ sc.doy + sc.doy2+ year +(1|date) , 
                  data=fishscapes.full, family=glmmTMB::lognormal(link="log"),
               ziformula=~ sc.doy + sc.doy2 +year+(1|date))

mod.lakeID.caughtBy.lmb<-glmmTMB(cpue~ sc.doy + sc.doy2+year+(1|lakeID)+(1|caughtBy), 
                  data=fishscapes.full, family=glmmTMB::lognormal(link="log"),
               ziformula=~ sc.doy + sc.doy2 +year+(1|lakeID)+(1|caughtBy) )

mod.lakeID.date.lmb<-glmmTMB(cpue~ sc.doy + sc.doy2+year+(1|lakeID)+(1|date) , 
                  data=fishscapes.full, family=glmmTMB::lognormal(link="log"),
               ziformula=~ sc.doy + sc.doy2 +year+(1|lakeID)+(1|date))

mod.caughtBy.date.lmb<-glmmTMB(cpue~ sc.doy + sc.doy2+year+(1|caughtBy)+(1|date) , 
                  data=fishscapes.full, family=glmmTMB::lognormal(link="log"),
               ziformula=~ sc.doy + sc.doy2+year+(1|caughtBy)+(1|date))
mod.lakeID.caughtBy.date.lmb<-glmmTMB(cpue~ sc.doy + sc.doy2+year+(1|lakeID)+(1|date)+(1|caughtBy) , 
                  data=fishscapes.full, family=glmmTMB::lognormal(link="log"),
               ziformula=~ sc.doy + sc.doy2 +year+(1|lakeID)+(1|date)+(1|caughtBy) )

AIC(mod.null.lmb, mod.lakeID.lmb, mod.caughtBy.lmb, mod.date.lmb, mod.lakeID.caughtBy.lmb, mod.lakeID.date.lmb, mod.caughtBy.date.lmb, mod.lakeID.caughtBy.date.lmb )


mod.fs.full<-mod.lakeID.caughtBy.date.lmb

mod.resid<-simulateResiduals(mod.fs.full, plot=T)
plot(mod.resid, form=fishscapes.full$sc.doy)
plot(mod.resid, form=fishscapes.full$sc.doy2)
plot(mod.resid, form=fishscapes.full$year)
plot(mod.resid, form=fishscapes.full$lakeID)
plot(mod.resid, form=fishscapes.full$date)
plot(mod.resid, form=fishscapes.full$caughtBy)

```

```{r}
summary(mod.fs.full)
performance::r2(mod.fs.full)
```










only random effects

```{r}
mod.random<-glmmTMB(cpue~(1|lakeID)+(1|date)+(1|caughtBy) , 
                  data=fishscapes.full, family=glmmTMB::lognormal(link="log"),
               ziformula=~(1|lakeID)+(1|date)+(1|caughtBy) )

simulateResiduals(mod.random, plot=T)
```
some metric of fish abundance is probably necessary. without it, I get these super weird residuals
