---
title: "nonlinear stan modeling"
output: html_document
date: "2024-04-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, here, lubridate, lme4, rstan, shinystan, truncnorm)
```

Fitting nonlinear models in Stan with fake data. 

This is a helpful tutorial using fishery data https://www.weirdfishes.blog/blog/fitting-bayesian-models-with-stan-and-r/ 


```{r}
catch.fun.err<-function(effort, q, popDensity, beta, error){
  catch=(effort*q*popDensity^beta)*error
  return(catch)
}

catch.fun<-function(effort, q, popDensity, beta){
  catch=(effort*q*popDensity^beta)
  return(catch)
}


# q as a truncated (at zero) normal? 

fake.df<-data.frame(popDensity=rlnorm(100, meanlog=2, sdlog=0.2),
                    effort=rlnorm(100, meanlog=1.25, sdlog=0.25),
                    q=rtruncnorm(100, a=0, b=Inf, mean=0.5, sd=1),
                    beta=rlnorm(100, meanlog=-1, 0.3),
                    error=rlnorm(100, meanlog=0, sdlog=0.3))

fake.df$catch<-floor(catch.fun(fake.df$effort, fake.df$q, fake.df$popDensity, fake.df$beta, fake.df$error))

fake.data<-list(N=nrow(fake.df),
                effort=fake.df$effort,
                lmbCatch=fake.df$catch,
                popDensity=fake.df$popDensity)
```
estimate q and beta
```{r}
fake_fit<-stan(file="fake.nonlinear.nb.stan",
                 data=fake.data,
                 chains=4,
                 warmup=1000,
                 iter=2000,
                 cores=4,
                 refresh=0)

print(fake_fit, pars=c("q","beta","phi", "lp__"), probs=c(.025,.5,.975))

traceplot(fake_fit, pars=c("q","beta","phi","lp__"), inc_warmup=FALSE, nrow=2)

```

Now try to add random effects. I'll make this more trivial by just repeating rows of the fake data
I want estimate a grand mean q and anglerID-specific q coefficients

left off here trying to parse example https://mc-stan.org/docs/stan-users-guide/regression.html#hierarchical-regression 
```{r}

fake.df<-data.frame(popDensity=rlnorm(100, meanlog=2, sdlog=0.2),
                    effort=rlnorm(100, meanlog=1.25, sdlog=0.25),
                    #q=rtruncnorm(100, a=0, b=Inf, mean=0.5, sd=1),
                    q=rlnorm(100, -0.5, 0.5),
                    beta=rlnorm(100, meanlog=-1, 0.3),
                    error=rlnorm(100, meanlog=0, sdlog=0.3))

fake.df$catch<-floor(catch.fun.err(fake.df$effort, fake.df$q, fake.df$popDensity, fake.df$beta, fake.df$error))

fake.df.rep<-fake.df%>%
  mutate(anglerID=seq(1:nrow(fake.df)))%>%
  group_by(anglerID)%>%
  slice(rep(1:n(), each=5))

fake.data.hier<-list(A=max(fake.df.rep$anglerID),
                     N=nrow(fake.df.rep),
                     AA=fake.df.rep$anglerID,
                     lmbCatch=fake.df.rep$catch,
                     effort=fake.df.rep$effort,
                     popDensity=fake.df.rep$popDensity
                     )

fake_fit_hier<-stan(file="fake.hierarchical.nonlinear.nb.stan",
                 data=fake.data.hier,
                 control=list(stepsize=1),
                 chains=4,
                 warmup=1000,
                 iter=2000,
                 cores=4,
                 refresh=0)

# print(fake_fit, pars=c("q","beta","phi", "lp__"), probs=c(.025,.5,.975))
# 
# traceplot(fake_fit, pars=c("q","beta","phi","lp__"), inc_warmup=FALSE, nrow=2)

launch_shinystan(fake_fit_hier)

```

Useful example in answers here of nonlinear hierarchical model https://stackoverflow.com/questions/50749414/developing-hierarchical-version-of-nonlinear-growth-curve-model-in-stan 