---
title: "prior checks"
output: html_document
date: "2024-08-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, here, lubridate, lme4, loo, rstan, shinystan, truncnorm, MFEUtilities, RColorBrewer, cowplot, bayesplot, cmdstanr, posterior, ggsci, cowplot)
```

Read in selected model fit, do some visualization on parameter estimates. Then revise model script (generated quantities block) to generate catch predictions for angler skill quantiles at mean popDensity and mean log_q_l, d. Finally, generate catch predictions across popDensities and dates

```{r}
rm(list=ls())
```

```{r}
# raw catch data
wd<-getwd()
db.dir<-paste0(wd, "/MFEdb/")
db<-"MFEdb_20220405.db"

dbTableList(db.dir, db)

lakes<-dbTable("lakes", fpath=db.dir, dbname=db)
#crew<-dbTable("crew", fpath=db.dir, dbname=db)%>%
#  filter(year%in%c("2018","2019"))
#write.csv(crew, "crew.csv")

# lake data

lakes.key<-lakes%>%
  dplyr::select(lakeID, lakeName, lat, long, WBIC, surfaceArea)
# Found lake missing surface area--135.97 ha
lakes.key[lakes.key$lakeID=="FD","surfaceArea"]<-135.97


projects<-dbTable("projects", fpath=db.dir, dbname=db)

fishSamples<-dbTable("fish_samples", fpath=db.dir, dbname=db)%>%
  filter(projectID%in%c("37") & gear=="AN")%>%
    mutate(nAnglers=as.numeric(nAnglers),
    effort=effort/nAnglers)


fishInfo<-dbTable("fish_info", fpath=db.dir, dbname=db)%>%
  filter(sampleID%in%fishSamples$sampleID)%>%
  filter(str_length(caughtBy)<4)%>%
  mutate(caughtBy=str_trim(caughtBy))%>%
  filter(otu=="largemouth_bass")%>%
  dplyr::select(projectID:caughtBy, comments)%>%
  # join fishing effort
  left_join(fishSamples[,c("lakeID","sampleID","dayOfYear","dateSample","dateTimeSample","crew","effort","effortUnits","nAnglers")], by="sampleID")

# catch rates for each angler trip, correcting for some ambiguous initials first

ALK.lake.date<-c("LV_20190608",
                 "SM_20190615",
                 "WN_20190617",
                 "BY_20190622")

AMK.lake.date<-c("DS_20190613",
                 "BOT_20190615",
                 "SM_20190621",
                 "BOT_20190624",
                 "NH_20190713")

long.crew<-fishSamples%>%
    mutate(date=str_split_fixed(sampleID, "_", 4)[,3],
         lakeID=str_split_fixed(sampleID, "_", 2)[,1],
         lakeID_date=paste(lakeID, date, sep="_"))%>%
  mutate(crewFix.ALK=ifelse(lakeID_date%in%ALK.lake.date & grepl("AK", crew), 1, 0),
         crewFix.AMK=ifelse(lakeID_date%in%AMK.lake.date & grepl("AK", crew), 1, 0))%>%
  mutate(crew=ifelse(crewFix.ALK==1, str_replace(crew, "AK", "ALK"), crew),
         crew=ifelse(crewFix.AMK==1, str_replace(crew, "AK", "AMK"), crew))%>%
  group_by(sampleID)%>%
  summarize(crew=unique(crew),
            effort=unique(effort))%>%
  # split crew into columns and then pivot longer
  separate("crew", paste("angler", 1:3, sep="_"), sep=", ", extra="drop")%>%
  pivot_longer(cols=angler_1:angler_3, names_to="angler_num", values_to="caughtBy", values_drop_na=T)%>%
  mutate(caughtBy=ifelse(caughtBy=="CMI","CI", caughtBy))%>%
  mutate(date=str_split_fixed(sampleID, "_", 4)[,3],
         lakeID=str_split_fixed(sampleID, "_", 2)[,1],
         lakeID_date=paste(lakeID, date, sep="_"))

 
 all.data<-fishInfo%>%
  mutate(year=year(dateSample))%>%
  left_join(lakes.key, by="lakeID")
  

  
# now get catch rates from angling data and left join to long.crew. empty values can then be replaced with 0

cpue<-all.data%>%
  group_by(sampleID, caughtBy)%>%
  summarize(nCaught=n())



full.data<-long.crew%>%
  left_join(cpue, by=c("sampleID", "caughtBy"))%>%
  mutate(nCaught=ifelse(is.na(nCaught), 0, nCaught))


```


code for processing mark recap data

```{r}
fish_samples<-dbTable("fish_samples", fpath=db.dir)%>%
  filter(projectID%in%c("37"))%>%
  filter(useSampleMarkRecap=="yes")

# ok, we want to do projects 37 and 38 separately. 37 used AF (anal fin) tags, 38 used PIT tags
fish_info<-dbTable("fish_info", fpath=db.dir)%>%
  filter(projectID%in%c("37"))%>%
  filter(sampleID%in%fish_samples$sampleID)%>%
  filter(otu=="largemouth_bass")

fish_data<-inner_join(fish_samples, fish_info, by="sampleID")%>%
  mutate(lakeID=str_split_fixed(sampleID, "_", 2)[,1],
         clipRecapture=as.numeric(clipRecapture),
         clipRecapture=ifelse(is.na(clipRecapture),0,clipRecapture),
         tagged=ifelse(clipApply=="AF", 1, 0))

samples<-unique(fish_data$sampleID)

lakes<-dbTable("lakes", fpath=db.dir)

fish_samples<-data.frame(sampleID=unique(fish_data$sampleID))%>%
  mutate(method=str_split_fixed(sampleID, "_", 6)[,5],
         sampleDate=str_split_fixed(sampleID, "_", 4)[,3],
         sampleTime=str_split_fixed(sampleID, "_", 5)[,4],
         date_time=ymd_hm(paste(sampleDate, sampleTime, sep="_")),
         # adjust sampleDates for night electrofishing--if method==BE and sampleTime is in the evening before midnight, add one day to date.
         adjust=ifelse(method=="BE" & sampleTime<2359 & sampleTime>1200, 1, 0),
         adj_sampleDate=as.character(ymd(sampleDate)+days(1)),
         batchDate=ifelse(adjust==1, adj_sampleDate, as.character(ymd(sampleDate))),
         batchDate_method=paste(batchDate, method, sep="_"))

fish_pe<-left_join(fish_data, fish_samples[,c("sampleID","batchDate_method")], by="sampleID")%>%
  group_by(lakeID, batchDate_method)%>%
  summarize(markedNow=sum(tagged, na.rm=T),
            recapturedNow=sum(clipRecapture))%>%
  mutate(marked_cum=cumsum(markedNow),
         markedPrior=lag(marked_cum),
         markedPrior=ifelse(is.na(markedPrior), 0, markedPrior),
         allFishCaught=markedNow+recapturedNow)%>%
  ungroup()

# let's filter to lakes that had at least 1 recapture

recap_count<-fish_pe%>%
  group_by(lakeID)%>%
  summarize(nRecap=sum(recapturedNow))%>%
  arrange(desc(nRecap))%>%
  filter(nRecap>0)

recap_over_1<-filter(recap_count, nRecap>1)

# surface area for Found lake was missing. Got it from WI DNR (find a lake tool) and converted to hectares
lakes[lakes$lakeID=="FD",]$surfaceArea<-136.0

fish_pe_recap<-fish_pe%>%
  filter(lakeID%in%recap_count$lakeID)%>%
  mutate(CtMt=allFishCaught*markedPrior)%>%
  group_by(lakeID)%>%
  summarize(sumCtMt=sum(CtMt),
            sumRt=sum(recapturedNow))%>%
  ungroup()%>%
  left_join(lakes[,c("lakeID","surfaceArea")], by="lakeID")

```

```{r}
data.join<-full.data%>%
  filter(lakeID%in%fish_pe_recap$lakeID)%>%
  left_join(fish_pe_recap, by="lakeID")


lakeID<-data.frame(lakeID=fish_pe_recap$lakeID,
                   L=seq(1:13),
                   sumCtMt=fish_pe_recap$sumCtMt,
                   sumRt=fish_pe_recap$sumRt,
                   surfaceArea=fish_pe_recap$surfaceArea)

anglerID<-data.frame(caughtBy=unique(data.join$caughtBy),
                     A=seq(1:18))

dateID<-data.frame(date=unique(data.join$date),
                   D=seq(1:42))

data.indexed<-data.join%>%
  left_join(dateID, by="date")%>%
  left_join(anglerID, by="caughtBy")%>%
  left_join(lakeID, by="lakeID")%>%
  mutate(log_effort=log(effort))%>%
  rename("lmbCatch"=nCaught)

z_scores <- c(1.645, 0.674, 0, -0.674, -1.645)

data.list<-list(N=nrow(data.indexed),
                A=max(data.indexed$A),
                D=max(data.indexed$D),
                L=max(data.indexed$L),
                AA=data.indexed$A,
                DD=data.indexed$D,
                LL=data.indexed$L,
                lmbCatch=data.indexed$lmbCatch,
                log_effort=data.indexed$log_effort,
                sumCtMt=lakeID$sumCtMt,
                sumRt=lakeID$sumRt,
                surfaceArea=lakeID$surfaceArea,
                z_scores = z_scores,
                z_scores_long= rep(z_scores, each=21),
                z_scores_best_worst_angler = rep(c(z_scores[1], z_scores[1], z_scores[5], z_scores[5]), each=21),
                z_scores_best_worst_date = rep(c(z_scores[1], z_scores[5], z_scores[1], z_scores[5]), each=21),
                z_scores_medium_angler = rep(c(z_scores[2], z_scores[2], z_scores[4], z_scores[4]), each=21),
                z_scores_medium_date = rep(c(z_scores[2], z_scores[4], z_scores[2], z_scores[4]), each=21),
                log_popDensity_pred=log(seq(1, 201, by=10))
                )


```


```{r}
set.seed(392)

# phi was the main problem

 inits<-function(){
   # narrower distribution for initial phi value (high values were breaking model fits in some chains)
   phi=rgamma(1,3)
   # beta=rlnorm(1,-1,1)
   # popDensity = rlnorm(1,0,2)
   # log_mu_q_a~rnorm(1,0,1)
   # log_mu_q_d~rnorm(1,0,1)
   # log_mu_q_l~rnorm(1,0,1)
   # q_a_raw~rnorm(1,0,1)
   # q_d_raw~rnorm(1,0,1)
   # q_l_raw~rnorm(1,0,1)
   # sigma_q_a~rexp(1,1)
   # sigma_q_d~rexp(1,1)
   # sigma_q_l~rexp(1,1)
 
   init<-list(phi=phi
              # beta=beta,
              # popDensity=popDensity,
              # # log_mu_q_a=log_mu_q_a,
              # # log_mu_q_d=log_mu_q_d,
              # # log_mu_q_l=log_mu_q_l,
              # q_a_raw=q_a_raw,
              # q_d_raw=q_d_raw,
              # q_l_raw=q_l_raw,
              # sigma_q_a=sigma_q_a,
              # sigma_q_d=sigma_q_d,
              # sigma_q_l=sigma_q_l
              )
   return(init)
 }

 fit_student<-stan(file=here("output_selected_model","prior_checks","models_prior_checks","prior_check_a_d_uninform_10.stan"),
                  data=data.list,
                  init=inits,
                  control=list(stepsize=0.1, max_treedepth=15),
                  chains=4,
                  warmup=1000,
                  iter=3000,
                  cores=4,
                  refresh=0)
 
 saveRDS(fit_student, here("output_selected_model","fits","student_t_priors_10.rds"))
#fit_student<-readRDS(here("output_selected_model","fits","student_t_priors_10.rds"))

post_student<-extract(fit_student)


```

```{r}
prop_zero<-function(x) mean(x==0)

log_lik_all_int<-extract_log_lik(fit_student, merge_chains=FALSE)
r_eff_all_int<-relative_eff(exp(log_lik_all_int))
loo_all_int<-loo(log_lik_all_int, r_eff=r_eff_all_int, cores=4, save_psis=TRUE)

post_student<-extract(fit_student)
ppc_dens_overlay(data.list$lmbCatch, post_student$posterior_pred_check[c(1:100),])+xlim(0,20)
#ggsave(here("output_selected_model", "posterior_predictive_checks", "student_t", "posterior_overlay_xlim_20.png"))

ppc_dens_overlay(data.list$lmbCatch, post_student$posterior_pred_check[c(1:100),])
#ggsave(here("output_selected_model", "posterior_predictive_checks", "student_t", "posterior_overlay_no_lim.png"))


ppc_hist(data.list$lmbCatch, post_student$posterior_pred_check[c(1:5),])
#ggsave(here("output_selected_model", "posterior_predictive_checks", "student_t", "posterior_hist_5_runs.png"))


ppc_stat(data.list$lmbCatch, post_student$posterior_pred_check, stat="prop_zero", binwidth=0.005)
#ggsave(here("output_selected_model", "posterior_predictive_checks", "student_t", "prop_zero_compare.png"))

ppc_stat(data.list$lmbCatch, post_student$posterior_pred_check, stat="max", binwidth=0.005)
#ggsave(here("output_selected_model", "posterior_predictive_checks", "student_t", "max_compare.png"))


ppc_dens_overlay_grouped(data.list$lmbCatch, post_student$posterior_pred_check[1:100,], group=data.list$AA)+xlim(0,10)
#ggsave(here("output_selected_model", "posterior_predictive_checks", "student_t", "posterior_overlay_angler_groups.png"))


ppc_dens_overlay_grouped(data.list$lmbCatch, post_student$posterior_pred_check[1:100,], group=data.list$LL)+xlim(0,20)
#ggsave(here("output_selected_model", "posterior_predictive_checks", "student_t", "posterior_overlay_lake_groups.png"))


ppc_dens_overlay_grouped(data.list$lmbCatch, post_student$posterior_pred_check[1:100,], group=data.list$DD)+xlim(0,10)
#ggsave(here("output_selected_model", "posterior_predictive_checks", "student_t", "posterior_overlay_date_groups.png"))


```

```{r}



beta<-post_student$beta
phi<-post_student$phi
sigma_q_a<-(post_student$sigma_q_a)
sigma_q_d<-(post_student$sigma_q_d)

sigma_e<-post_student$log_resid_sd
sigma_fixed<-post_student$sigma_fixed
sigma_total<-post_student$sigma_total

sigma.total.df<-data.frame(param=c(rep("sigma_total", length(sigma_total))),
                           sd=sigma_total)

sigma.fixed.df<-data.frame(param=c(rep("sigma_fixed", length(sigma_fixed))),
                           sd=sigma_fixed)

sigma.error.df<-data.frame(param=c(rep("sigma_error", length(sigma_e))),
                           sd=sigma_e)

sigma.df<-data.frame(param=c(rep("sigma_q_a", length(sigma_q_a)),
                             rep("sigma_q_d", length(sigma_q_d))),
                     sd=c(sigma_q_a, sigma_q_d))

popDensity<-as.data.frame(post_student$popDensity)

log_q_mu<-post_student$log_q_mu

log_q_a<-as.data.frame(post_student$log_q_a)
log_q_d<-as.data.frame(post_student$log_q_d)

names_a<-paste0("angler_", seq(1:18))
names_d<-as.character(unique(data.indexed$date))
names_l<-paste0("lake_", seq(1:13))

names(log_q_a)<-names_a
names(log_q_d)<-names_d
#names(log_q_l)<-names_l


# log.q.df<-data.frame(param=c(rep("log_q_mu", length(log_q_mu)),
#                              rep("log_q_a", length(log_q_a)),
#                              rep("log_q_d", length(log_q_d)),
#                              rep("log_q_l", length(log_q_l))),
#                      estimate=c(log_q_mu, log_q_a, log_q_d, log_q_l))


angler_long<-log_q_a%>%
  pivot_longer(cols=everything(), names_to="angler", values_to="angler_effect")%>%
  mutate(angler=factor(angler, levels=c("angler_1","angler_2","angler_3","angler_4","angler_5","angler_6","angler_7","angler_8","angler_9","angler_10","angler_11","angler_12","angler_13","angler_14","angler_15","angler_16","angler_17","angler_18")))

date_long<-log_q_d%>%
  pivot_longer(cols=everything(), names_to="date", values_to="date_effect")%>%
  mutate(date=as.factor(as.numeric(date)))
  


```

density plot figures 

fun color palettes

```{r}
library(scales)
futurama<-pal_futurama("planetexpress")(12)
simpsons<-pal_simpsons("springfield")(12)

```

```{r}

beta.df<-data.frame(param=rep("beta"), estimate=beta)
```

Compare priors to posteriors
For first round, set scale parameter of student t to 1. (popDensity was 50) 
On next round, set parameter to 10. these are saved with _10 at tthe end. I changed the popDensity scale prior from 50 to 100 and phi gamma to 1, 0.1 (from 1, 0.5)
```{r}
popDensity.prior.10<-post_student$prior_t_popDensity[post_student$prior_t_popDensity>=0]
half.t.prior.10<-post_student$prior_t_other[post_student$prior_t_other>=0]
phi.prior.10<-post_student$prior_phi
prior.sigmas.10<-post_student$prior_sigma

beta.prior.10<-half.t.prior.10
phi.prior.10<-phi.prior.10
sigma.q.priors.10<-half.t.prior.10

post<-post_student

popDensity.post<-post$popDensity
beta.post<-post$beta
phi.post<-post$phi
sigma.q.a.post<-post$sigma_q_a
sigma.q.d.post<-post$sigma_q_d
sigma.q.l.post<-post$sigma_q_l

popDensity.post.df<-as.data.frame(popDensity.post)
names(popDensity.post.df)<-paste0("lake_", seq(1:13))

popDensity.plot<-popDensity.post.df%>%
  pivot_longer(cols=everything(), names_to="lake", values_to="popDensity")%>%
  mutate(lake=factor(lake, levels=c("lake_1","lake_2","lake_3","lake_4","lake_5","lake_6","lake_7","lake_8","lake_9","lake_10",
                                    "lake_11","lake_12","lake_13")))

popDensity.100<-ggplot()+
  geom_density(data=popDensity.plot, aes(x=popDensity, fill=lake),bounds=c(0, Inf), alpha=0.5)+
    geom_density(aes(x=popDensity.prior.10),bounds=c(0, Inf), fill="gray", alpha=0.7)+
  xlim(0,200)+
  xlab("Population density (1/ha)")+
   ylab(expression(paste("Probability density p(", "Population density", "|Catch)")))+
  theme_bw()
popDensity.100
ggsave(here("output_selected_model", "prior_checks", "prior_checks_student","popDensity_prior_check_100.png"))
```


```{r}
beta.10<-ggplot()+
  geom_density(aes(x=beta.prior.10), bounds=c(0, Inf),fill="gray")+
  geom_density(aes(x=beta.post), bounds=c(0, Inf), fill="lightblue")+
  xlab(expression(beta))+
   ylab(expression(paste("Probability density p(", beta, "|Catch)")))+

  xlim(0,10)+
  theme_bw()
beta.10
  ggsave(here("output_selected_model", "prior_checks", "prior_checks_student", "beta_prior_check_10.png"))

```

```{r}
phi.10<-ggplot()+
  geom_density(aes(x=phi.prior.10), bounds=c(0, Inf), fill="gray")+
  geom_density(aes(x=phi.post),bounds=c(0, Inf), fill="lightblue")+
  ylab(expression(paste("Probability density p(", phi, "|Catch)")))+
  xlab(expression(phi))+
  xlim(0,55)+
  theme_bw()
phi.10
ggsave(here("output_selected_model", "prior_checks", "prior_checks_student","phi_prior_check_10.png"))

```
Phi is pretty sensitive to prior 

```{r}
sigma.check.10<-data.frame(group=c(rep("prior", length(sigma.q.priors.10)), rep("angler", 8000), rep("date", 8000)),
                        sigma.q=c(sigma.q.priors.10, sigma.q.a.post, sigma.q.d.post))

sigma.10<-ggplot(sigma.check.10)+
  geom_density(aes(x=sigma.check.10$sigma.q, fill=group),bounds=c(0, Inf), alpha=0.5)+
  scale_fill_manual("Effect",values=c("lightpink","lightyellow","gray"))+
  xlim(0,10)+
   ylab(expression(paste("Probability density p(", sigma["q"], "|Catch)")))+
  xlab(expression(sigma["q"]))+
  theme_bw()
sigma.10
ggsave(here("output_selected_model", "prior_checks", "prior_checks_student","sigma_prior_check_10.png"))

```
Everything except phi is very similar with a wide, very vague prior. Exception: the poor population estimates are even worse, but this does not result in changes to the variance partitioning or beta estimate. The lake effect ICC posterior does get wider, suggesting it's starting to take up some of that variation unexplained by pop density.


```{r}
set.seed(482)
fit_student_1<-stan(file=here("output_selected_model","prior_checks","models_prior_checks","prior_check_a_d_uninform_1.stan"),
                  data=data.list,
                  init=inits,
                  control=list(stepsize=0.1, max_treedepth=15),
                  chains=4,
                  warmup=1000,
                  iter=3000,
                  cores=4,
                  refresh=0)
 
 saveRDS(fit_student_1, here("output_selected_model","fits","student_t_priors_1.rds"))
#fit_student_1<-readRDS(here("output_selected_model","fits","student_t_priors_1.rds"))

post_student_1<-extract(fit_student_1)

```


```{r}



beta<-post_student_1$beta
phi<-post_student_1$phi
sigma_q_a<-(post_student_1$sigma_q_a)
sigma_q_d<-(post_student_1$sigma_q_d)

sigma_e<-post_student_1$log_resid_sd
sigma_fixed<-post_student_1$sigma_fixed
sigma_total<-post_student_1$sigma_total

sigma.total.df<-data.frame(param=c(rep("sigma_total", length(sigma_total))),
                           sd=sigma_total)

sigma.fixed.df<-data.frame(param=c(rep("sigma_fixed", length(sigma_fixed))),
                           sd=sigma_fixed)

sigma.error.df<-data.frame(param=c(rep("sigma_error", length(sigma_e))),
                           sd=sigma_e)

sigma.df<-data.frame(param=c(rep("sigma_q_a", length(sigma_q_a)),
                             rep("sigma_q_d", length(sigma_q_d)),
                     sd=c(sigma_q_a, sigma_q_d)))

popDensity<-as.data.frame(post_student_1$popDensity)

log_q_mu<-post_student_1$log_q_mu

log_q_a<-as.data.frame(post_student_1$log_q_a)
log_q_d<-as.data.frame(post_student_1$log_q_d)

names_a<-paste0("angler_", seq(1:18))
names_d<-as.character(unique(data.indexed$date))
names_l<-paste0("lake_", seq(1:13))

names(log_q_a)<-names_a
names(log_q_d)<-names_d


# log.q.df<-data.frame(param=c(rep("log_q_mu", length(log_q_mu)),
#                              rep("log_q_a", length(log_q_a)),
#                              rep("log_q_d", length(log_q_d)),
#                              rep("log_q_l", length(log_q_l))),
#                      estimate=c(log_q_mu, log_q_a, log_q_d, log_q_l))


angler_long<-log_q_a%>%
  pivot_longer(cols=everything(), names_to="angler", values_to="angler_effect")%>%
  mutate(angler=factor(angler, levels=c("angler_1","angler_2","angler_3","angler_4","angler_5","angler_6","angler_7","angler_8","angler_9","angler_10","angler_11","angler_12","angler_13","angler_14","angler_15","angler_16","angler_17","angler_18")))

date_long<-log_q_d%>%
  pivot_longer(cols=everything(), names_to="date", values_to="date_effect")%>%
  mutate(date=as.factor(as.numeric(date)))
  


```


Compare priors to posteriors
For first round, set scale parameter of student t to 1. (popDensity was 50) (Note: first round I didnt' change sigma prior; I need to go back and do this)
On next round, set parameter to 10. these are saved with _10 at tthe end. I changed the popDensity scale prior from 50 to 100
```{r}
popDensity.prior<-post_student_1$prior_t_popDensity[post_student_1$prior_t_popDensity>=0]
half.t.prior<-post_student_1$prior_t_other[post_student_1$prior_t_other>=0]
phi.prior<-post_student_1$prior_phi
prior.sigma<-post_student_1$prior_sigma

beta.prior<-half.t.prior
phi.prior<-phi.prior
sigma.q.priors<-half.t.prior

post<-post_student_1

popDensity.post<-post$popDensity
beta.post<-post$beta
phi.post<-post$phi
sigma.q.a.post<-post$sigma_q_a
sigma.q.d.post<-post$sigma_q_d

popDensity.post.df<-as.data.frame(popDensity.post)
names(popDensity.post.df)<-paste0("lake_", seq(1:13))

popDensity.plot<-popDensity.post.df%>%
  pivot_longer(cols=everything(), names_to="lake", values_to="popDensity")%>%
  mutate(lake=factor(lake, levels=c("lake_1","lake_2","lake_3","lake_4","lake_5","lake_6","lake_7","lake_8","lake_9","lake_10",
                                    "lake_11","lake_12","lake_13")))

popDensity.50<-ggplot()+
  geom_density(data=popDensity.plot, aes(x=popDensity, fill=lake), bounds=c(0, Inf), alpha=0.5)+
    geom_density(aes(x=popDensity.prior), bounds=c(0, Inf), fill="gray", alpha=0.7)+
  xlim(0,200)+
  xlab("Population density (1/ha)")+
   ylab(expression(paste("Probability density p(", "Population density", "|Catch)")))+
  theme_bw()
popDensity.50

ggsave(here("output_selected_model", "prior_checks", "prior_checks_student","popDensity_prior_check_50.png"))
```


```{r}
beta.1<-ggplot()+
  geom_density(aes(x=beta.prior), bounds=c(0, Inf),fill="gray")+
  geom_density(aes(x=beta.post), bounds=c(0, Inf), fill="lightblue")+
  xlab(expression(beta))+
   ylab(expression(paste("Probability density p(", beta, "|Catch)")))+

  xlim(0,10)+
  theme_bw()
beta.1
  ggsave(here("output_selected_model", "prior_checks", "prior_checks_student", "beta_prior_check_1.png"))

```

```{r}
phi.1<-ggplot()+
  geom_density(aes(x=phi.prior), bounds=c(0, Inf),fill="gray")+
  geom_density(aes(x=phi.post),bounds=c(0, Inf),  fill="lightblue")+
  ylab(expression(paste("Probability density p(", phi, "|Catch)")))+
  xlab(expression(phi))+
  xlim(0,55)+
  theme_bw()
phi.1
ggsave(here("output_selected_model", "prior_checks", "prior_checks_student","phi_prior_check_1.png"))

```

```{r}

sigma.check<-data.frame(group=c(rep("prior", length(sigma.q.priors)), rep("angler", 8000), rep("date", 8000)),
                        sigma.q=c(sigma.q.priors, sigma.q.a.post, sigma.q.d.post))

sigma.1<-ggplot(sigma.check)+
  geom_density(aes(x=sigma.check$sigma.q, fill=group), bounds=c(0, Inf), alpha=0.5)+
  scale_fill_manual("Effect",values=c("lightpink","lightyellow","gray"))+
  xlim(0,10)+
   ylab(expression(paste("Probability density p(", sigma["q"], "|Catch)")))+
  xlab(expression(sigma["q"]))+
  theme_bw()
sigma.1

ggsave(here("output_selected_model", "prior_checks", "prior_checks_student","sigma_prior_check_1.png"))

```
comparison plots

```{r}
library(cowplot)

plot_grid(popDensity.50, popDensity.100, rel_widths=c(1, 1.3), labels=c("A","B"))
ggsave(here("output_selected_model","prior_checks","prior_checks_student", "popDensity.compare.png"), height=4, width=8)

```

