---
title: "model_summary_visualization"
output: html_document
date: "2024-08-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, here, lubridate, lme4, loo, rstan, shinystan, truncnorm, MFEUtilities, RColorBrewer, cowplot, bayesplot, cmdstanr, posterior, ggsci, cowplot)
```

Read in selected model fit, do some visualization on parameter estimates. Then revise model script (generated quantities block) to generate catch predictions for angler skill quantiles at mean popDensity and mean log_q_l, d. Finally, generate catch predictions across popDensities and dates

```{r}
rm(list=ls())
```

```{r}
# raw catch data
wd<-getwd()
db.dir<-paste0(wd, "/MFEdb/")
db<-"MFEdb_20220405.db"

dbTableList(db.dir, db)

lakes<-dbTable("lakes", fpath=db.dir, dbname=db)
#crew<-dbTable("crew", fpath=db.dir, dbname=db)%>%
#  filter(year%in%c("2018","2019"))
#write.csv(crew, "crew.csv")

# lake data

lakes.key<-lakes%>%
  dplyr::select(lakeID, lakeName, lat, long, WBIC, surfaceArea)
# Found lake missing surface area--135.97 ha
lakes.key[lakes.key$lakeID=="FD","surfaceArea"]<-135.97


projects<-dbTable("projects", fpath=db.dir, dbname=db)

fishSamples<-dbTable("fish_samples", fpath=db.dir, dbname=db)%>%
  filter(projectID%in%c("37") & gear=="AN")%>%
    mutate(nAnglers=as.numeric(nAnglers),
    effort=effort/nAnglers)


fishInfo<-dbTable("fish_info", fpath=db.dir, dbname=db)%>%
  filter(sampleID%in%fishSamples$sampleID)%>%
  filter(str_length(caughtBy)<4)%>%
  mutate(caughtBy=str_trim(caughtBy))%>%
  filter(otu=="largemouth_bass")%>%
  dplyr::select(projectID:caughtBy, comments)%>%
  # join fishing effort
  left_join(fishSamples[,c("lakeID","sampleID","dayOfYear","dateSample","dateTimeSample","crew","effort","effortUnits","nAnglers")], by="sampleID")

# catch rates for each angler trip, correcting for some ambiguous initials first

ALK.lake.date<-c("LV_20190608",
                 "SM_20190615",
                 "WN_20190617",
                 "BY_20190622")

AMK.lake.date<-c("DS_20190613",
                 "BOT_20190615",
                 "SM_20190621",
                 "BOT_20190624",
                 "NH_20190713")

long.crew<-fishSamples%>%
    mutate(date=str_split_fixed(sampleID, "_", 4)[,3],
         lakeID=str_split_fixed(sampleID, "_", 2)[,1],
         lakeID_date=paste(lakeID, date, sep="_"))%>%
  mutate(crewFix.ALK=ifelse(lakeID_date%in%ALK.lake.date & grepl("AK", crew), 1, 0),
         crewFix.AMK=ifelse(lakeID_date%in%AMK.lake.date & grepl("AK", crew), 1, 0))%>%
  mutate(crew=ifelse(crewFix.ALK==1, str_replace(crew, "AK", "ALK"), crew),
         crew=ifelse(crewFix.AMK==1, str_replace(crew, "AK", "AMK"), crew))%>%
  group_by(sampleID)%>%
  summarize(crew=unique(crew),
            effort=unique(effort))%>%
  # split crew into columns and then pivot longer
  separate("crew", paste("angler", 1:3, sep="_"), sep=", ", extra="drop")%>%
  pivot_longer(cols=angler_1:angler_3, names_to="angler_num", values_to="caughtBy", values_drop_na=T)%>%
  mutate(caughtBy=ifelse(caughtBy=="CMI","CI", caughtBy))%>%
  mutate(date=str_split_fixed(sampleID, "_", 4)[,3],
         lakeID=str_split_fixed(sampleID, "_", 2)[,1],
         lakeID_date=paste(lakeID, date, sep="_"))

 
 all.data<-fishInfo%>%
  mutate(year=year(dateSample))%>%
  left_join(lakes.key, by="lakeID")
  

  
# now get catch rates from angling data and left join to long.crew. empty values can then be replaced with 0

cpue<-all.data%>%
  group_by(sampleID, caughtBy)%>%
  summarize(nCaught=n())



full.data<-long.crew%>%
  left_join(cpue, by=c("sampleID", "caughtBy"))%>%
  mutate(nCaught=ifelse(is.na(nCaught), 0, nCaught))


```


code for processing mark recap data

```{r}
fish_samples<-dbTable("fish_samples", fpath=db.dir)%>%
  filter(projectID%in%c("37"))%>%
  filter(useSampleMarkRecap=="yes")

# ok, we want to do projects 37 and 38 separately. 37 used AF (anal fin) tags, 38 used PIT tags
fish_info<-dbTable("fish_info", fpath=db.dir)%>%
  filter(projectID%in%c("37"))%>%
  filter(sampleID%in%fish_samples$sampleID)%>%
  filter(otu=="largemouth_bass")

fish_data<-inner_join(fish_samples, fish_info, by="sampleID")%>%
  mutate(lakeID=str_split_fixed(sampleID, "_", 2)[,1],
         clipRecapture=as.numeric(clipRecapture),
         clipRecapture=ifelse(is.na(clipRecapture),0,clipRecapture),
         tagged=ifelse(clipApply=="AF", 1, 0))

samples<-unique(fish_data$sampleID)

lakes<-dbTable("lakes", fpath=db.dir)

fish_samples<-data.frame(sampleID=unique(fish_data$sampleID))%>%
  mutate(method=str_split_fixed(sampleID, "_", 6)[,5],
         sampleDate=str_split_fixed(sampleID, "_", 4)[,3],
         sampleTime=str_split_fixed(sampleID, "_", 5)[,4],
         date_time=ymd_hm(paste(sampleDate, sampleTime, sep="_")),
         # adjust sampleDates for night electrofishing--if method==BE and sampleTime is in the evening before midnight, add one day to date.
         adjust=ifelse(method=="BE" & sampleTime<2359 & sampleTime>1200, 1, 0),
         adj_sampleDate=as.character(ymd(sampleDate)+days(1)),
         batchDate=ifelse(adjust==1, adj_sampleDate, as.character(ymd(sampleDate))),
         batchDate_method=paste(batchDate, method, sep="_"))

fish_pe<-left_join(fish_data, fish_samples[,c("sampleID","batchDate_method")], by="sampleID")%>%
  group_by(lakeID, batchDate_method)%>%
  summarize(markedNow=sum(tagged, na.rm=T),
            recapturedNow=sum(clipRecapture))%>%
  mutate(marked_cum=cumsum(markedNow),
         markedPrior=lag(marked_cum),
         markedPrior=ifelse(is.na(markedPrior), 0, markedPrior),
         allFishCaught=markedNow+recapturedNow)%>%
  ungroup()

# let's filter to lakes that had at least 1 recapture

recap_count<-fish_pe%>%
  group_by(lakeID)%>%
  summarize(nRecap=sum(recapturedNow))%>%
  arrange(desc(nRecap))%>%
  filter(nRecap>0)

recap_over_1<-filter(recap_count, nRecap>1)

# surface area for Found lake was missing. Got it from WI DNR (find a lake tool) and converted to hectares
lakes[lakes$lakeID=="FD",]$surfaceArea<-136.0

fish_pe_recap<-fish_pe%>%
  filter(lakeID%in%recap_count$lakeID)%>%
  mutate(CtMt=allFishCaught*markedPrior)%>%
  group_by(lakeID)%>%
  summarize(sumCtMt=sum(CtMt),
            sumRt=sum(recapturedNow))%>%
  ungroup()%>%
  left_join(lakes[,c("lakeID","surfaceArea")], by="lakeID")


fish_pe_for_table<-fish_pe%>%
  filter(lakeID%in%recap_count$lakeID)%>%
  mutate(CtMt=allFishCaught*markedPrior)%>%
  group_by(lakeID)%>%
  summarize(sumCtMt=sum(CtMt),
            sumRt=sum(recapturedNow),
            sumCaught=sum(allFishCaught),
            nSampleEvents=n())%>%
  ungroup()%>%
  left_join(lakes[,c("lakeID","surfaceArea")], by="lakeID")%>%
    select(lakeID, surfaceArea, sumCaught, sumRt, nSampleEvents, sumCtMt)


```

```{r}
data.join<-full.data%>%
  filter(lakeID%in%fish_pe_recap$lakeID)%>%
  left_join(fish_pe_recap, by="lakeID")


lakeID<-data.frame(lakeID=fish_pe_recap$lakeID,
                   L=seq(1:13),
                   sumCtMt=fish_pe_recap$sumCtMt,
                   sumRt=fish_pe_recap$sumRt,
                   surfaceArea=fish_pe_recap$surfaceArea)

anglerID<-data.frame(caughtBy=unique(data.join$caughtBy),
                     A=seq(1:18))

dateID<-data.frame(date=unique(data.join$date),
                   D=seq(1:42))

data.indexed<-data.join%>%
  left_join(dateID, by="date")%>%
  left_join(anglerID, by="caughtBy")%>%
  left_join(lakeID, by="lakeID")%>%
  mutate(log_effort=log(effort))%>%
  rename("lmbCatch"=nCaught)

z_scores <- c(1.645, 0.674, 0, -0.674, -1.645)

data.list<-list(N=nrow(data.indexed),
                A=max(data.indexed$A),
                D=max(data.indexed$D),
                L=max(data.indexed$L),
                AA=data.indexed$A,
                DD=data.indexed$D,
                LL=data.indexed$L,
                lmbCatch=data.indexed$lmbCatch,
                log_effort=data.indexed$log_effort,
                sumCtMt=lakeID$sumCtMt,
                sumRt=lakeID$sumRt,
                surfaceArea=lakeID$surfaceArea,
                z_scores = z_scores,
                z_scores_long= rep(z_scores, each=21),
                z_scores_best_worst_angler = rep(c(z_scores[1], z_scores[1], z_scores[5], z_scores[5]), each=21),
                z_scores_best_worst_date = rep(c(z_scores[1], z_scores[5], z_scores[1], z_scores[5]), each=21),
                z_scores_medium_angler = rep(c(z_scores[2], z_scores[2], z_scores[4], z_scores[4]), each=21),
                z_scores_medium_date = rep(c(z_scores[2], z_scores[4], z_scores[2], z_scores[4]), each=21),
                log_popDensity_pred=log(seq(1, 201, by=10))
                )


```

Lake table--name, WBIC, number (L), location, surface area, n observations, n marks, n recaps

```{r}
lakeTable.fs<-read_csv(here("data","lakeTable.update.3.csv"))

units<-dbTable("units", fpath=db.dir, dbname=db)%>%
  filter(tableName=="LAKES" & colName=="surfaceArea") 
# yes, in hectares

bay<-projects<-dbTable("lakes", fpath=db.dir, dbname=db)%>%
  select(lakeID, lakeName, surfaceArea, lat, long, WBIC)%>%
  filter(lakeID=="BA")


n.obs<-data.indexed%>%
  group_by(lakeID)%>%
  # note that multiple entries were sometimes listed per angler effort block, so nDates*nAnglers!=nObs
  summarize(nObs=n(),
            nDates=length(unique(date)),
            nAnglers=length(unique(caughtBy)))%>%
  left_join(lakeTable.fs[c("lakeID","lakeName","WBIC","lat","long")], by="lakeID")

bay.row<-c("BA",5,2,4,bay$lakeName, NA,  bay$lat, bay$long)

lake.table.n<-rbind.data.frame(n.obs[1,], bay.row, n.obs[3:13,])%>%
  left_join(lakeID[,c("lakeID", "L")], by="lakeID")%>%
  left_join(fish_pe_for_table, by="lakeID")%>%
  select(lakeName, L, WBIC, lat, long, surfaceArea ,nObs, nDates, nAnglers, 
         sumCaught, sumRt, nSampleEvents, sumCtMt)%>%
  rename("nAnglingObs"=nObs,
         "nAnglingDates"=nDates,
         "totalCaught"=sumCaught,
         "totalRecaptured"=sumRt)
write.csv(lake.table.n, here("output_selected_model","tables","lake.table.csv"))
```


```{r}
#fit<-readRDS(here("candidate_models","candidate_model_fits","fit_all_int.rds"))
```

```{r}
set.seed(26)

# phi was the main problem

 inits<-function(){
   # narrower distribution for initial phi value (high values were breaking model fits in some chains)
   phi=rgamma(1,3)
   # beta=rlnorm(1,-1,1)
   # popDensity = rlnorm(1,0,2)
   # log_mu_q_a~rnorm(1,0,1)
   # log_mu_q_d~rnorm(1,0,1)
   # log_mu_q_l~rnorm(1,0,1)
   # q_a_raw~rnorm(1,0,1)
   # q_d_raw~rnorm(1,0,1)
   # q_l_raw~rnorm(1,0,1)
   # sigma_q_a~rexp(1,1)
   # sigma_q_d~rexp(1,1)
   # sigma_q_l~rexp(1,1)
 
   init<-list(phi=phi
              # beta=beta,
              # popDensity=popDensity,
              # # log_mu_q_a=log_mu_q_a,
              # # log_mu_q_d=log_mu_q_d,
              # # log_mu_q_l=log_mu_q_l,
              # q_a_raw=q_a_raw,
              # q_d_raw=q_d_raw,
              # q_l_raw=q_l_raw,
              # sigma_q_a=sigma_q_a,
              # sigma_q_d=sigma_q_d,
              # sigma_q_l=sigma_q_l
              )
   return(init)
 }

fit_gen<-stan(file=here("candidate_models","selected_model_all_int.stan"),
                 data=data.list,
                 #init=inits,
                 control=list(stepsize=0.1, max_treedepth=15),
                 chains=4,
                 warmup=1000,
                 iter=3000,
                 cores=4,
                 refresh=0)
saveRDS(fit_gen, here("output_selected_model","fits","model_with_predictions.rds"))

post_gen<-extract(fit_gen)


```

summary table
```{r}

fit.table<-rstan::summary(fit_gen, pars=c("beta","phi","sigma_q_a","sigma_q_d","sigma_q_l", "popDensity",
                               "log_q_a","log_q_d","log_q_l",
                               "log_q_mu","log_mu_q_a","log_mu_q_d","log_mu_q_l","q_a_raw",
                         "q_d_raw","q_l_raw","lp__"))

table.summary<-fit.table$summary
write.csv(table.summary, here::here("output_selected_model","tables","fit_all_int_param_summary.csv"))
```


First thing is a nice set of plots showing the posteriors of beta, each sigma, the mean intercept, and the angler-, lake-, and date-specific intercepts

```{r}

beta<-post_gen$beta
phi<-post_gen$phi
sigma_q_a<-(post_gen$sigma_q_a)
sigma_q_d<-(post_gen$sigma_q_d)
sigma_q_l<-(post_gen$sigma_q_l)

sigma_random <-post_gen$sigma_random

sigma_fixed<-post_gen$sigma_fixed
sigma_pred<-post_gen$sigma_pred

sigma.pred.df<-data.frame(param=c(rep("sigma_predictions", length(sigma_pred))),
                           sd=sigma_pred)

sigma.fixed.df<-data.frame(param=c(rep("sigma_fixed", length(sigma_fixed))),
                           sd=sigma_fixed)

sigma.random.df<-data.frame(param=c(rep("sigma_random", length(sigma_random))),
                           sd=sigma_random)

sigma.df<-data.frame(param=c(rep("sigma_q_a", length(sigma_q_a)),
                             rep("sigma_q_d", length(sigma_q_d)),
                             rep("sigma_q_l", length(sigma_q_l))),
                     sd=c(sigma_q_a, sigma_q_d, sigma_q_l))

popDensity<-as.data.frame(post_gen$popDensity)

log_q_mu<-post_gen$log_q_mu

log_q_a<-as.data.frame(post_gen$log_q_a)
log_q_d<-as.data.frame(post_gen$log_q_d)
log_q_l<-as.data.frame(post_gen$log_q_l)

names_a<-paste0("angler_", seq(1:18))
names_d<-as.character(unique(data.indexed$date))
names_l<-paste0("lake_", seq(1:13))

names(log_q_a)<-names_a
names(log_q_d)<-names_d
names(log_q_l)<-names_l


# log.q.df<-data.frame(param=c(rep("log_q_mu", length(log_q_mu)),
#                              rep("log_q_a", length(log_q_a)),
#                              rep("log_q_d", length(log_q_d)),
#                              rep("log_q_l", length(log_q_l))),
#                      estimate=c(log_q_mu, log_q_a, log_q_d, log_q_l))


angler_long<-log_q_a%>%
  pivot_longer(cols=everything(), names_to="angler", values_to="angler_effect")%>%
  mutate(angler=factor(angler, levels=c("angler_1","angler_2","angler_3","angler_4","angler_5","angler_6","angler_7","angler_8","angler_9","angler_10","angler_11","angler_12","angler_13","angler_14","angler_15","angler_16","angler_17","angler_18")))

date_long<-log_q_d%>%
  pivot_longer(cols=everything(), names_to="date", values_to="date_effect")%>%
  mutate(date=as.factor(as.numeric(date)))
  
lake_long<-log_q_l%>%
  pivot_longer(cols=everything(), names_to="lake", values_to="lake_effect")%>%
  mutate(lake=factor(lake, levels=c("lake_1","lake_2","lake_3","lake_4","lake_5","lake_6","lake_7","lake_8","lake_9","lake_10","lake_11","lake_12","lake_13")))


```

density plot figures 

fun color palettes

```{r}
library(scales)
futurama<-pal_futurama("planetexpress")(12)
simpsons<-pal_simpsons("springfield")(12)

```

compare posteriors with priors

```{r}
popDensity.prior<-post_gen$prior_t_popDensity[post_gen$prior_t_popDensity>=0]
half.t.prior<-post_gen$prior_t_other[post_gen$prior_t_other>=0]
phi.prior<-post_gen$prior_phi
prior.sigma<-half.t.prior

beta.prior<-half.t.prior
sigma.q.priors<-half.t.prior

post<-post_gen

popDensity.post<-post$popDensity
beta.post<-post$beta
phi.post<-post$phi
sigma.q.a.post<-post$sigma_q_a
sigma.q.d.post<-post$sigma_q_d
sigma.q.l.post<-post$sigma_q_l

popDensity.post.df<-as.data.frame(popDensity.post)
names(popDensity.post.df)<-paste0("lake_", seq(1:13))

popDensity.plot<-popDensity.post.df%>%
  pivot_longer(cols=everything(), names_to="lake", values_to="popDensity")

ggplot()+
  geom_density(data=popDensity.plot, aes(x=popDensity, fill=lake), alpha=0.5)+
    geom_density(aes(x=popDensity.prior), fill="gray", alpha=0.7,  bounds=c(0, Inf))+
  xlim(0,200)+
  theme_bw()

ggsave(here("output_selected_model", "figures", "popDensity_prior_check.png"))
```


```{r}
ggplot()+
  geom_density(aes(x=beta.prior), bounds=c(0, Inf), fill="gray")+
  geom_density(aes(x=beta.post), fill="lightblue")+
  ggtitle("beta")+
  xlim(0,10)+
  theme_bw()
  ggsave(here("output_selected_model", "figures",  "beta_prior_check.png"))

```

```{r}
ggplot()+
  geom_density(aes(x=phi.prior), fill="gray",  bounds=c(0, Inf))+
  geom_density(aes(x=phi.post), fill="lightblue")+
  ggtitle("phi")+
  xlim(0,15)+
  theme_bw()
ggsave(here("output_selected_model", "figures", "phi_prior_check.png"))

```

```{r}
ggplot()+
  geom_density(aes(x=sigma.q.priors), fill="gray",  bounds=c(0, Inf))+
  geom_density(aes(x=sigma.q.a.post), fill="lightyellow", alpha=0.5)+
  geom_density(aes(x=sigma.q.d.post), fill="lightgreen", alpha=0.5)+
  geom_density(aes(x=sigma.q.l.post), fill="lightpink", alpha=0.5)+
  ggtitle("sigma.q.a.yellow.d.green.l.red")+
  xlim(0,10)+
  theme_bw()
ggsave(here("output_selected_model", "figures","sigma_prior_check.png"))

```



```{r}

beta.df<-data.frame(param=rep("beta"), estimate=beta)


sigma.prediction.plot<-ggplot(sigma.pred.df)+
  geom_density(aes(x=sd, fill=param), alpha=0.7)+
  geom_vline(xintercept=sd(data.list$lmbCatch), linetype="dashed")+
  scale_fill_manual(values=futurama[2], name="", labels=expression(sigma[hat(Catch)]))+
  xlim(0,8)+
  xlab(expression(sigma))+
  ylab(expression(paste("p(", sigma[hat(Catch)], "|Catch)")))+
  theme_bw()+
  theme(legend.text=element_text(size=16),
        axis.title.y=element_text(size=14),
        axis.title.x=element_text(size=14))


# sigma.e.plot<-ggplot(sigma.error.df)+
#   geom_density(aes(x=sd, fill=param), alpha=0.7)+
#   scale_fill_manual(values=futurama[5], name="", labels=expression(sigma[epsilon]))+
#   xlim(0,7)+
#   xlab(expression(sigma[epsilon]))+
#   ylab(expression(paste("p(", sigma[epsilon], "|catch)")))+
#   theme_bw()+
#   theme(legend.text=element_text(size=16),
#         axis.title.y=element_text(size=14),
#         axis.title.x=element_blank())
# sigma.e.plot

sigma.fixed.plot<-ggplot(sigma.fixed.df)+
  geom_density(aes(x=sd, fill=param), alpha=0.7)+
  scale_fill_manual(values=futurama[6], name="", labels=expression(sigma[beta]))+
  xlim(0,8)+
  xlab(expression(sigma[beta]))+
  ylab(expression(paste("p(", sigma[beta], "|catch)")))+
  theme_bw()+
  theme(legend.text=element_text(size=16),
        axis.title.y=element_text(size=14),
        axis.title.x=element_blank())

 # beta.plot<-ggplot(beta.df)+
 #   geom_density(aes(x=estimate, fill=param), alpha=0.7)+
 #   scale_fill_manual(values=futurama[2], name="", labels=expression(beta))+
 #   xlim(0,7)+
 #   xlab(expression(beta))+
 #   ylab(expression(paste("p(", beta, "|catch)")))+
 #   theme_bw()
 # beta.plot

sigma.random.plot<-ggplot(sigma.df)+
  geom_density(aes(x=sd, fill=param), alpha=0.7)+
  scale_fill_manual(values=futurama[c(1,4,3)], name="", labels=expression(sigma[a], sigma[d], sigma[l]))+
  xlim(0,8)+
  ylab(expression(paste("p(", sigma["a,d,l"], "|catch)")))+
  xlab(expression(sigma))+
  theme_bw()+
  theme(legend.text=element_text(size=16),
        axis.title.y=element_text(size=14),
        axis.title.x=element_text(size=16))
sigma.random.plot

plot_grid(sigma.random.plot, sigma.fixed.plot,  sigma.prediction.plot, ncol=1, align="v")
#plot_grid(sigma.total.plot, sigma.e.plot, sigma.fixed.plot, sigma.random.plot, ncol=1, align="v")
gsave(here("output_selected_model","figures","beta.sigma.posterior.png"), height=8, width=6)
```


```{r}
obs<-data.list$lmbCatch
pred<-post_gen$posterior_pred_check[50,]

ggplot()+
  geom_point(aes(x=obs, y=pred))+
  geom_abline(slope=1, intercept=0, linetype="dashed")+
  ylim(0,40)+
  xlim(0,40)+
  theme_bw()
```



```{r}
angler_long_plot<-angler_long%>%
  group_by(angler)%>%
  summarize(mean=mean(angler_effect), 
            sd=sd(angler_effect),
            upper95=mean+1.96*sd,
            lower95=mean-1.96*sd)%>%
  ungroup()%>%
  mutate(angler=as.factor(as.numeric(str_replace(angler, "angler_", ""))))

angler<-ggplot(angler_long_plot)+
  geom_point(aes(x=mean, y=angler))+
  geom_errorbarh(aes(y=angler, xmin=lower95, xmax=upper95))+
  ylab("Angler observed")+
  xlab(expression(paste("Angler skill effect (ln", q["a"], ")")))+
  theme_bw()

  
  angler
  ggsave(here("output_selected_model","figures","angler.effects.png"), height=4, width=6)
```

```{r}
date_long_plot<-date_long%>%
  group_by(date)%>%
  summarize(mean=mean(date_effect), 
            sd=sd(date_effect),
            upper95=mean+1.96*sd,
            lower95=mean-1.96*sd)%>%
  ungroup()%>%
  mutate(date=as.factor(ymd(date)))

date<-ggplot(date_long_plot)+
  geom_point(aes(x=mean, y=date))+
  geom_errorbarh(aes(y=date, xmin=lower95, xmax=upper95))+
  ylab("Date observed")+
  xlab(expression(paste("Daily conditions effect (ln", q["d"], ")")))+
  theme_bw()

  
  date
  ggsave(here("output_selected_model","figures","date.effects.png"), height=6, width=6)

```

Lake effect

```{r}
lake_long_plot<-lake_long%>%
  group_by(lake)%>%
  summarize(mean=mean(lake_effect), 
            sd=sd(lake_effect),
            upper95=mean+1.96*sd,
            lower95=mean-1.96*sd)%>%
  ungroup()%>%
  mutate(lake=as.factor(as.numeric(str_replace(lake, "lake_", ""))))

lake<-ggplot(lake_long_plot)+
  geom_point(aes(x=mean, y=lake))+
  geom_errorbarh(aes(y=lake, xmin=lower95, xmax=upper95))+
  ylab("Lake observed")+
  xlab(expression(paste("Lake effect (ln", q["l"], ")")))+
  theme_bw()

  
  lake
    ggsave(here("output_selected_model","figures","lake.effects.png"), height=4, width=6)


```
Now predictions of mean catch for each angler--are violin plots really appropriate since they're discrete predictions?

```{r}
angler.pred<-as.data.frame(post_gen$predict_angler_catch)

names(angler.pred)<-paste0("angler_", seq(1:18))

angler.pred.long<-angler.pred%>%
  pivot_longer(cols=everything(), names_to="angler", values_to="catch")%>%
  mutate(angler=as.factor(as.numeric(str_replace(angler, "angler_", ""))))

ggplot(angler.pred.long)+
  geom_violin(aes(x=angler, y=catch), fill=futurama[7])+
  theme_bw()+
  theme(legend.position="none")+
  coord_flip()+
  xlab("Angler observed")+
  ylab("Predicted mean catch")
ggsave(here::here("output_selected_model","figures","angler.mean.pred.catch.png"), height=6, width=4)
```
That's not very useful for readers; instead do quantiles

```{r}
quant.catch.df<-as.data.frame(post_gen$predict_quantile_catch)

names(quant.catch.df)<-c("95th", "75th", "50th", "25th", "5th")

quant.catch.long<-quant.catch.df%>%
  pivot_longer(cols=everything(), names_to="Angler skill", values_to="Mean catch")%>%
  mutate(`Angler skill`=factor(`Angler skill`, levels=c("5th", "25th", "50th", "75th", "95th")))

quant.catch.sum<-quant.catch.long%>%
  group_by(`Angler skill`)%>%
  summarize(mean=mean(`Mean catch`),
            median=median(`Mean catch`),
            sd=sd(`Mean catch`),
            upper=mean+1.96*sd,
            lower=mean-1.96*sd)

ggplot(quant.catch.long)+
  geom_violin(aes(x=`Angler skill`, y=`Mean catch`), fill=futurama[1], adjust=2)+
  geom_point(data=quant.catch.sum, aes(x=`Angler skill`, y=mean), size=3)+
  geom_point(data=quant.catch.sum, aes(x=`Angler skill`, y=median), size=3, color="gray")+
  coord_flip()+
  xlab(expression(paste("Angler catch power percentile (ln ", "q"["a"], ")")))+
  ylab("Predicted mean catch")+
  theme_bw()+
  theme(axis.title.y=element_text(size=16),
        axis.text.y=element_text(size=12),
        axis.text.x=element_text(size=12),
        axis.title.x=element_text(size=16))

ggsave(here("output_selected_model","figures", "percentile.catch.violin.png"), height=4, width=6)
```

Finally for these quantiles of anglers across a range of population densities

Showing here: there is a lot of unexplained variation, yes. The mean effect of angler skill and daily conditions, however, is still in the range that's noticeable

55 rows of predictions
```{r}
popDensity<-rep(seq(1,201, by=10), 5)
percentile<-rep(c("95%","75%","50%","25%","5%"), each=21)


angler.pred.pop<-post_gen$predict_quantile_angler_catch_popDens

catch.angler.df<-data.frame(popDensity=popDensity,
                            percentile=factor(percentile, levels=rev(c("95%","75%","50%","25%","5%"))),
                            meanCatch=colMeans(angler.pred.pop))

gradient.a<-c("#ffdce2","#ffced1","#ffafa6","#ff8d6b","#FF6F00FF")

ggplot(catch.angler.df)+
  geom_line(aes(x=popDensity, y=meanCatch, color=percentile), linewidth=2)+
  scale_color_manual(values=gradient.a)+
  guides(color=guide_legend(title=expression(paste("Angler skill\neffect percentile"))))+
  xlab("Population density (1/ha)")+
  ylab("Predicted mean catch")+
  theme_bw()+
    theme(axis.title.y=element_text(size=16),
        axis.text.y=element_text(size=12),
        axis.text.x=element_text(size=12),
        axis.title.x=element_text(size=16))

ggsave(here("output_selected_model","figures","angler.effect.popDensity.png"), height=4, width=6)
```
Next daily conditions

```{r}
quant.date.catch.df<-as.data.frame(post_gen$predict_quantile_date_catch)

names(quant.date.catch.df)<-c("95th", "75th", "50th", "25th", "5th")

quant.date.catch.long<-quant.date.catch.df%>%
  pivot_longer(cols=everything(), names_to="Day effect", values_to="Mean catch")%>%
  mutate(`Day effect`=factor(`Day effect`, levels=c("5th", "25th", "50th", "75th", "95th")))

quant.date.catch.sum<-quant.date.catch.long%>%
  group_by(`Day effect`)%>%
  summarize(mean=mean(`Mean catch`),
            median=median(`Mean catch`),
            sd=sd(`Mean catch`),
            upper=mean+1.96*sd,
            lower=mean-1.96*sd)

ggplot(quant.date.catch.long)+
  geom_violin(aes(x=`Day effect`, y=`Mean catch`), fill="#8A4198FF", adjust=2)+
  geom_point(data=quant.date.catch.sum, aes(x=`Day effect`, y=mean), size=3)+
  geom_point(data=quant.date.catch.sum, aes(x=`Day effect`, y=median), size=3, color="gray")+
  coord_flip()+
  xlab(expression(paste("Day effect percentile (ln ", "q"["d"], ")")))+
  ylab("Predicted mean catch")+
  theme_bw()+
  theme(axis.title.y=element_text(size=16),
        axis.text.y=element_text(size=12),
        axis.text.x=element_text(size=12),
        axis.title.x=element_text(size=16))

ggsave(here("output_selected_model","figures", "percentile.day.catch.violin.png"), height=4, width=6)

```


```{r}
popDensity<-rep(seq(1,201, by=10), 5)
percentile<-rep(c("95%","75%","50%","25%","5%"), each=21)


date.pred.pop<-post_gen$predict_quantile_date_catch_popDens

catch.date.df<-data.frame(popDensity=popDensity,
                            percentile=factor(percentile, levels=rev(c("95%","75%","50%","25%","5%"))),
                            meanCatch=colMeans(date.pred.pop))

gradient.a<-c("#efd2dd","#d9a6c3","#bf7caf","#9e55a0","#8A4198FF")

ggplot(catch.date.df)+
  geom_line(aes(x=popDensity, y=meanCatch, color=percentile), linewidth=2)+
  scale_color_manual(values=gradient.a)+
  guides(color=guide_legend(title=expression(paste("Daily conditions\neffect percentile"))))+
  xlab("Population density (1/ha)")+
  ylab("Predicted mean catch")+
  theme_bw()+
    theme(axis.title.y=element_text(size=16),
        axis.text.y=element_text(size=12),
        axis.text.x=element_text(size=12),
        axis.title.x=element_text(size=16))

ggsave(here("output_selected_model","figures","date.effect.popDensity.png"), height=4, width=6)


```
As we'd expect from the sigma_d distribution, it's very similar

Effects together? highest percentiles and lowest percentiles


```{r}
show_col(simpsons)
c( "#F05C3BFF", "#197EC0FF")
```


```{r}
quantiles<-c("95%", "5%")


best.worst.df<-data.frame(pred.catch=colMeans(post_gen$predict_best_worst),
                          pop.density=rep(seq(1, 201, by=10), 4),
                          quant.angler=factor(rep(quantiles, each=42), levels=c("5%","95%")),
                          quant.date=factor(rep(quantiles, each=21), levels=c("5%","95%")))%>%
  mutate(group=ifelse(quant.angler=="95%" & quant.date=="95%", "High skill,\ngood conditions",
                      ifelse(quant.angler=="95%" & quant.date=="5%", "High skill,\npoor conditions",
                             ifelse(quant.angler=="5%" & quant.date=="95%", "Low skill,\ngood conditions",
                                    ifelse(quant.angler=="5%" & quant.date=="5%", "Low skill,\npoor conditions", "error")))),
         facet=ifelse(quant.angler=="95%", "High skill (95th percentile)", "Low skill (5th percentile)"),
         legend=ifelse(quant.date=="95%", "Good (95th percentile)", "Poor (5th percentile)"))

ggplot(best.worst.df)+
  geom_line(aes(x=pop.density, y=pred.catch, color=legend), linewidth=2)+
  scale_color_manual(values=c( "#F05C3BFF", "#197EC0FF"))+
  xlim(0,150)+
  facet_grid(.~facet)+
  guides(color=guide_legend(title="Daily conditions"))+
  xlab("Population density (1/ha)")+
  ylab("Predicted mean catch")+
  theme_bw()+
    theme(axis.title.y=element_text(size=16),
        axis.text.y=element_text(size=12),
        axis.text.x=element_text(size=12),
        axis.title.x=element_text(size=16),
        strip.text.x=element_text(size=11),
        legend.text=element_text(size=11),
        legend.title=element_text(size=12))
ggsave(here("output_selected_model","figures","best.worst.comparison.png"), height=4, width=8)

```

Ah, right, it's a multiplicative effect of angler skill and daily conditions. 

I'd like a version with 75-25 because these extremes are wild. 

```{r}
quantiles<-c("75%", "25%")


medium.df<-data.frame(pred.catch=colMeans(post_gen$predict_medium),
                          pop.density=rep(seq(1, 201, by=10), 4),
                          quant.angler=factor(rep(quantiles, each=42), levels=c("25%","75%")),
                          quant.date=factor(rep(quantiles, each=21), levels=c("25%","75%")))%>%
  mutate(facet=ifelse(quant.angler=="75%", "Higher skill (75th percentile)", "Lower skill (25th percentile)"),
         legend=ifelse(quant.date=="75%", "Good (75th percentile)", "Poor (25th percentile)"))

ggplot(medium.df)+
  geom_line(aes(x=pop.density, y=pred.catch, color=legend), linewidth=2)+
  scale_color_manual(values=c( "#F05C3BFF", "#197EC0FF"))+
  xlim(0,150)+
  facet_grid(.~facet)+
  guides(color=guide_legend(title="Daily conditions"))+
  xlab("Population density (1/ha)")+
  ylab("Predicted mean catch")+
  theme_bw()+
    theme(axis.title.y=element_text(size=16),
        axis.text.y=element_text(size=12),
        axis.text.x=element_text(size=12),
        axis.title.x=element_text(size=16),
        strip.text.x=element_text(size=11),
        legend.text=element_text(size=11),
        legend.title=element_text(size=12))
ggsave(here("output_selected_model","figures","medium.comparison.png"), height=4, width=8)

```

Compare to angler date model
```{r}
set.seed(9287)


fit_red<-stan(file=here("candidate_models","selected_model_angler_date_int.stan"),
                 data=data.list,
                 #init=inits,
                 control=list(stepsize=0.1, max_treedepth=15),
                 chains=4,
                 warmup=1000,
                 iter=3000,
                 cores=4,
                 refresh=0)
saveRDS(fit_red, here("output_selected_model","fits","reduced_model_with_predictions.rds"))

post_red<-extract(fit_red)


```

table

```{r}
fit.red.table<-rstan::summary(fit_red, pars=c("beta","phi","sigma_q_a","sigma_q_d","popDensity",
                               "log_q_a","log_q_d",
                               "log_q_mu","log_mu_q_a","log_mu_q_d","q_a_raw",
                         "q_d_raw","lp__"))

table.red.summary<-fit.red.table$summary
write.csv(table.red.summary, here::here("output_selected_model","tables","fit_angler_date_int_param_summary.csv"))

```

First thing is a nice set of plots showing the posteriors of beta, each sigma, the mean intercept, and the angler-, lake-, and date-specific intercepts

```{r}

beta<-post_red$beta
phi<-post_red$phi
sigma_q_a<-(post_red$sigma_q_a)
sigma_q_d<-(post_red$sigma_q_d)

sigma_e<-post_red$log_resid_sd
sigma_fixed<-post_red$sigma_fixed
sigma_total<-post_red$sigma_total

sigma.total.df<-data.frame(param=c(rep("sigma_total", length(sigma_total))),
                           sd=sigma_total)

sigma.fixed.df<-data.frame(param=c(rep("sigma_fixed", length(sigma_fixed))),
                           sd=sigma_fixed)

sigma.error.df<-data.frame(param=c(rep("sigma_error", length(sigma_e))),
                           sd=sigma_e)

sigma.df<-data.frame(param=c(rep("sigma_q_a", length(sigma_q_a)),
                             rep("sigma_q_d", length(sigma_q_d))),
                     sd=c(sigma_q_a, sigma_q_d))

popDensity<-as.data.frame(post_red$popDensity)

log_q_mu<-post_red$log_q_mu

log_q_a<-as.data.frame(post_red$log_q_a)
log_q_d<-as.data.frame(post_red$log_q_d)

names_a<-paste0("angler_", seq(1:18))
names_d<-as.character(unique(data.indexed$date))

names(log_q_a)<-names_a
names(log_q_d)<-names_d


# log.q.df<-data.frame(param=c(rep("log_q_mu", length(log_q_mu)),
#                              rep("log_q_a", length(log_q_a)),
#                              rep("log_q_d", length(log_q_d)),
#                              rep("log_q_l", length(log_q_l))),
#                      estimate=c(log_q_mu, log_q_a, log_q_d, log_q_l))


angler_long<-log_q_a%>%
  pivot_longer(cols=everything(), names_to="angler", values_to="angler_effect")%>%
  mutate(angler=factor(angler, levels=c("angler_1","angler_2","angler_3","angler_4","angler_5","angler_6","angler_7","angler_8","angler_9","angler_10","angler_11","angler_12","angler_13","angler_14","angler_15","angler_16","angler_17","angler_18")))

date_long<-log_q_d%>%
  pivot_longer(cols=everything(), names_to="date", values_to="date_effect")%>%
  mutate(date=as.factor(as.numeric(date)))
  


```



```{r}

beta.df<-data.frame(param=rep("beta"), estimate=beta)


sigma.total.plot<-ggplot(sigma.total.df)+
  geom_density(aes(x=sd, fill=param), alpha=0.7)+
  scale_fill_manual(values=futurama[2], name="", labels=expression(sigma["total"]))+
  xlim(0,7)+
  xlab(expression(sigma["total"]))+
  ylab(expression(paste("p(", sigma["total"], "|catch)")))+
  theme_bw()+
  theme(legend.text=element_text(size=16),
        axis.title.y=element_text(size=14),
        axis.title.x=element_blank())


sigma.e.plot<-ggplot(sigma.error.df)+
  geom_density(aes(x=sd, fill=param), alpha=0.7)+
  scale_fill_manual(values=futurama[5], name="", labels=expression(sigma[epsilon]))+
  xlim(0,7)+
  xlab(expression(sigma[epsilon]))+
  ylab(expression(paste("p(", sigma[epsilon], "|catch)")))+
  theme_bw()+
  theme(legend.text=element_text(size=16),
        axis.title.y=element_text(size=14),
        axis.title.x=element_blank())
sigma.e.plot

sigma.fixed.plot<-ggplot(sigma.fixed.df)+
  geom_density(aes(x=sd, fill=param), alpha=0.7)+
  scale_fill_manual(values=futurama[6], name="", labels=expression(sigma[beta]))+
  xlim(0,7)+
  xlab(expression(sigma[beta]))+
  ylab(expression(paste("p(", sigma[beta], "|catch)")))+
  theme_bw()+
  theme(legend.text=element_text(size=16),
        axis.title.y=element_text(size=14),
        axis.title.x=element_blank())

# beta.plot<-ggplot(beta.df)+
#   geom_density(aes(x=estimate, fill=param), alpha=0.7)+
#   scale_fill_manual(values=futurama[2], name="", labels=expression(beta))+
#   xlim(0,7)+
#   xlab(expression(beta))+
#   ylab(expression(paste("p(", beta, "|catch)")))+
#   theme_bw()
# beta.plot

sigma.random.plot<-ggplot(sigma.df)+
  geom_density(aes(x=sd, fill=param), alpha=0.7)+
  scale_fill_manual(values=futurama[c(1,4,3)], name="", labels=expression(sigma[a], sigma[d], sigma[l]))+
  xlim(0,7)+
  ylab(expression(paste("p(", sigma["a,d,l"], "|catch)")))+
  xlab(expression(sigma))+
  theme_bw()+
  theme(legend.text=element_text(size=16),
        axis.title.y=element_text(size=14),
        axis.title.x=element_text(size=16))
sigma.random.plot

plot_grid(sigma.random.plot, sigma.fixed.plot, sigma.e.plot, sigma.total.plot, ncol=1, align="v")
#plot_grid(sigma.total.plot, sigma.e.plot, sigma.fixed.plot, sigma.random.plot, ncol=1, align="v")
ggsave(here("output_selected_model","figures","beta.sigma.post.reduced.png"), height=8, width=6)
```
ICC
```{r}
icc.df<-data.frame(group=c(rep("Angler effect", 8000), rep("Date effect", 8000)),
                     ICC=c(post_red$ICC_a, post_red$ICC_d))

ggplot(icc.df)+
  geom_density(aes(x=ICC, fill=group), alpha=0.7)+
  scale_fill_manual(values=futurama[c(1,4,3)], name="")+
  ylab(expression(paste("p(", ICC["a,d,l"], "|catch)")))+
  xlab("Intraclass Correlation Coefficient")+
  theme_bw()+
    theme(legend.text=element_text(size=10),
        axis.title.y=element_text(size=12),
        axis.title.x=element_text(size=12))
ggsave(here::here("output_selected_model","figures","ICC_density.post.reduced.png"), height=4, width=6)


```
```{r}
r2.df<-data.frame(type=c(rep("marginal", 8000), rep("conditional", 8000)),
                  r2=c(post_red$r2_marginal, post_red$r2_conditional))

ggplot(r2.df)+
  geom_density(aes(x=r2, fill=type), alpha=0.7)+
  scale_fill_manual(values=simpsons[c(1,2)], name="")+
  ylab(expression(paste("p(", R^2, "|catch)")))+
  xlab(expression(R^2))+
  theme_bw()+
    theme(legend.text=element_text(size=10),
        axis.title.y=element_text(size=12),
        axis.title.x=element_text(size=12))
ggsave(here::here("output_selected_model","figures","R2_density.post.reduced.png"), height=4, width=6)

```
